Small Data Center using Raspberry Pi 2 for Video Streaming

### 2015-06-08 paper/paper.tex
paper changes

\usepackage[numbers]{natbib}
\setlength{\bibsep}{0.0pt}
\newcommand\todo[1]{
	\textcolor{green}{#1}
}
\restylefloat{figure}
\restylefloat{table}
\begin{document}
\setcounter{page}{1}
\pagenumbering{arabic}
%
% --- Author Metadata here --- DO NOT REMOVE OR CHANGE 
\conferenceinfo{23$^{th}$ Twente Student Conference on IT}{June 22$^{st}$, 2015, Enschede, The Netherlands.}
\CopyrightYear{2015} % Allows default copyright year (200X) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

\keywords{Cloud computing, Raspberry Pi 2,  micro data center, video streaming, load balancing}

\section{Introduction}
Today, most of us have some data in the cloud. Despite the attention from the community, research and development of Cloud Computing services are still in its early days~\citep{tso:2013}. \newline
Cloud computing is a trend in IT that moves customers computing and data away from computer and laptop into large data centers \citep{dikaiakos:2009}. In the future, most internet users are expected to access internet services over lightweight portable devices requiring a lot of data bandwidth \cite{dikaiakos:2009}. This bandwidth usage causes bottlenecks. To prevent these bottlenecks, data centers are build around the globe. These data centers can be greatly improved \cite{abrahamsson:2013,beloglazov:2010}. 
For example, these data centers require a lot of space and cooling. There are now new technologies such as for example the powerful ARM processor. Many companies want to explore the possibilities of the Raspberry Pi 2 , because of its ARM processor and the low price \cite{Pcextreme}. At the time of writing the Raspberry Pi 2 costs around 40 euro. Some data centers already offer some cloud computing using the Raspberry Pi 2 \cite{Pcextreme}. \newline
A Raspberry Pi 2 is a small computer and has a power usage between the 2 and 4 Watt. In contrast, a normal server has a power usage between the 75 and 250 watt~\cite{Pcextreme,beloglazov2012energy}. The low power consumption of the Raspberry Pi 2 and its computing power could mean that it is better to use a Raspberry Pi 2 for specific small tasks that do not demand an entire server. For this reason research is done on the performance of the Raspberry Pi 2 in a small data center. \newline
A Raspberry Pi cloud can be the small data center for the future \cite{tso:2013}. The Raspberry Pi 2 is cheaper in price compared to a more expensive normal server. The Raspberry Pi 2 can be an ideal testbed for testing distributed software.

In The Netherlands, one million people use Netflix \cite{volkskrant}. Netflix is a popular video streaming service that provides HD movies. For this Netflix makes use of a Content Distribution Network (CDN). Netflix makes use of MPEG-DASH a protocol for streaming over HTTP \cite{martin:2013}. The problem is that Netflix is responsible for  29,7\% of the peak downstream traffic in the US \cite{Adhikari:2012}, because of the traffic the two main providers Comcast and  AT\&T were limiting the downstream of Netflix. This caused a lot of criticism, to solve this a new law for net neutrality has been made \cite{net-neutrality}. \newline
In the Netherlands, video streaming services such as Youtube and Netflix have an increasing number of users. Nowadays, the data streaming problems are increasing, this makes it interesting to do more research in data centers with video streaming. Most of the existing research in computer science is performed on expensive large servers \cite{tso:2013}. It can be very useful to see if it is possible to do research on a small computer like a Raspberry Pi 2.

This paper investigates how useful the Raspberry Pi 2 is in a data center with video streaming. To come to a conclusion the performance of the Raspberry Pi 2 is evaluated by doing several benchmark tests. For this the main research question is: 
\begin{center}
How well does the Raspberry Pi 2 perform in small scale data centers with video streaming? 
\end{center}

@@

\begin{enumerate}
	\item What are small scale data centers with video streaming and why are they used?
	\item Is the Raspberry Pi 2 usable for video streaming?
	\item How to fit the Raspberry Pi's into a data center considering cooling, space allocation and power?
	\item What setup does a Raspberry Pi cloud with video streaming require?
	\item How is availability in a Raspberry pi cluster with video streaming affected by various load balancing techniques? 
\end{enumerate}

The five sub-questions are looked at in the next sections. Followed by a final section for conclusions and identified areas for future work.
The first two research question are a literature study into video streaming and cloud computing. and a part of it is  The last three sub-questions go into the technical research that investigates the performance and possibility of Raspberry Pi's for cloud video streaming. This is done by building a small Raspberry Pi cloud.

\section{Research Methods}
In the research the  aim is to build a Raspberry Pi cloud to measure the performance. This is achieved by building a Raspberry Pi cloud and doing benchmark test on this cloud. This research uses the Design Science method proposed by Hevner~\cite{hevner:2007}. The method is used to make an architecture for a Raspberry Pi cloud. This design science method consists of three cycles:
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{Design_science.png}
	\caption{Design Science cycles \cite{hevner:2007}}
	\label{fig:design} %always place your label after your caption!
\end{figure}
\textbf{relevance cycle} Represents the interaction with the domain. Here the problems and opportunities are identified. The identification of these problems are done in the next section. \newline 
\textbf{design cycle} The most important cycle where the design is build. This is highlighted in section \ref{sec:setupex}. \newline
\textbf{rigor cycle} Represents the interaction with the scientific community. This information can be found in section \ref{sec:back} and \ref{sec:setupex}.

\section{Background}\label{sec:back}
\subsection{Small scale computing}
Small scale cloud computing is cloud computing with smaller computation amounts than normal \cite{cox:2014}. Most customers for cloud solutions buy the services instead of buying the hardware and having to set up the servers themselves. This service is called Infrastructure as a Service (IaaS). The elasticity of computing resources in the cloud is important to provide IaaS \cite{Miettinen:2010:EEM:1863103.1863107}. Many companies have their software running on virtual machines in data centers \cite{beloglazov:2010}. The advantage is that these virtual machines are scaled dynamically to the customer its need. Using the cloud makes  management easier.

Several cloud services have a huge network demand, this may cause data centers to be unreachable. Because of the privacy and reachability of a data center many companies have their own servers. These servers are not energy efficient. \newline
Netflix a cloud service is responsible for approximately 30\% of the peak downstream traffic in US \cite{Adhikari:2012, computer-networking}. The data flowing through the network has different patterns. To improve this there are several data patterns possible, it can for example make a huge different in sending data in small bursts or in one large burst \cite{computer-networking}.

Power Usage Efficiency (PUE) is a important aspect in cloud computing \cite{beloglazov2012energy}. There are more lightweight portable devices coming \cite{Miettinen:2010:EEM:1863103.1863107}, these devices are energy efficient, but will demand more processes being done by data centers \cite{Miettinen:2010:EEM:1863103.1863107}. Nowadays, the data centers operate 12 to 20\% of the time, the other 80 \% of the time the servers inside are still using energy \cite{beloglazov2012energy}. Therefore there is a need Green Cloud Computing solutions that reduce the electrical power consumption and the environmental impact of data centers~\cite{beloglazov2012energy}. Microservers can be used to reduce the consumption.

\subsection{Microserver}
Microservers are used in data centers \cite{microserver}. Microservers consume less electricity and have enough computer power for small tasks, therefore they can be used for high traffic usage tasks. Besides this they are small and are easier replaced compared to a normal server, by using them data centers can be more dynamic \cite{microserver}. The ability to add and remove microservers more easily can save a lot of energy. \newline
A microserver often has a different processor compared to a normal server. The processors are often microprocessors like a ARM processor \cite{microserver}. There are two big processor companies in the world called ARM, Intel and AMD. The Intel and AMD processors are better at complex tasks, while the ARM processor is better at small tasks \cite{microserver}. In the data center of the future more dedicated microservers will be used, they are better at specialized tasks and are more energy efficient \cite{microserver}.

\subsection{Video streaming}\label{sec:stream}
Video streaming happens with HTTP, HTTP sends frames to the client and in the frames is the video stored \cite{computer-networking}. If the send rate is large enough the video will play without any delays, but if the send rate is too less then the video play will alternate between periods of continuous playout and periods of freezing~\cite{computer-networking}. \newline 
To make video streaming possible the client will connect with video streaming service its load balancer. The load balancer will then locate the videos. To show how the processes and operators operate with each other a UML sequence diagram is used \cite{Bernardi:2002:USD:584369.584376}. 
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{VideoStreaming.jpg}
	\caption{Video stream}
	\label{fig:stream} %always place your label after your caption!
\end{figure}

In the diagram there is a loop for the video streaming, here the client will keep requesting for more parts of the videos. Depending on the algorithm being used the amount of data requested by the client can fluctuate. For streaming a video HTTP streaming is often used. By using HTTP the client can correctly receive the video, if it receives too much data its buffer for video playing gets full. To prevent throwing data away the server then stops sending.

There is a important streaming technique called adaptive streaming. This streaming algorithm is often used for video on demand (VOD). It improves the experience for the end-user by adapting the video quality dynamically to the viewers network condition \cite{ffmpeg}. This techniques looks to the clients CPU possibility and its network possibilities. It delivers small sized chunks at multiple bitrates. This techniques is used for HTTP Live Streaming (HLS). This makes use of RTMP streaming and HTTP. This way of streaming is released in 2014. It codifies video in the H.264 format.\newline
There is another special kind of HTTP streaming called DASH (Dynamic Adaptive Streaming over HTTP)~\cite{computer-networking}. In DASH the video is encoded into several different versions. If the bandwidth is high, then the client selects chunks from a high-rate version and the way around~\cite{computer-networking}. The possibility to switch between bit rate versions is in video streaming very important. To make a smooth transition between these versions possible there are intermediate versions so that the client will see less suddenly the change in video quality. In HTTP streaming the server stores the video with a different URL.

\subsection{Video streaming and load balancing}
For video streaming several load balancing techniques are used. To choose a technique there is looked at performance, reliability and features. There can be a non-uniform demand for different movies. One solution to overcome this problem would be to replicate the most popular movies, but this is expensive in terms of storage space required for this. A dynamic replication can be used to solve the load demand~\cite{dan1996load}. By using this  portions of a movie are moved to less used storage devices, so that these can stream the movie. The allocation of these movies only happen when there is a to high increase on one specific storage device. In this way load imbalances can be prevented \cite{dan1996load}. \newline
Load balancers are generally grouped in two layers. Layer 4 (FTP,IP, UDP, TCP )and layer 7 the application layer. Layer 7 load balancers distribute request based upon data in the application layer. Here below are some protocols and languages used for this:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
	\item HTTP, is the hypertext transfer protocol. In section \ref{sec:stream} are mentioned some usages of this protocol. A lot of video streaming services use HTTP streaming \cite{Adhikari:2012}.  
	\item RTMP is a real time messaging protocol ~\cite{rtmp}. It is developed for streaming data between a player and a server. It is encapsulated in HTTP to traverse firewalls. It gives more video player options compared to normal HTTP thus giving the user a better experience. A disadvantage compared to normal HTTP is that it is sensitive for data spikes. These data spikes can result in a overload of the buffer which results in a empty buffer, this can lead to a stop in the video playing. 
	\item Synchronized Multimedia Integration Language (SMIL) describes multimedia presentations \cite{smil}. SMIL refers to media objects by URLs, allowing them to be shared between presentations and stored on different servers for load balancing. 
\end{enumerate}

@@

One thing that users want is on demand video streaming. They do not want to store the data itself, and they want to have a wide choice of different videos when they want to watch it. Netflix is a on demand video streaming service that makes HD movies watching possible. For this Netflix makes use of a content distribution network (CDN). They make use of amazon its AWS, simpleDB , S3 and Cassandra for file storage \cite{Adhikari:2012}. Netflix makes use of MPEG-DASH a protocol that makes streaming over HTTP possible. To make this possible Netflix does several things in the cloud.

@@

\subsubsection{Video stream CDN}
A CDN makes use of small data centers. This is because users want their data as nearby as possible  In this way the latency is decreased and there is a shorter connection distance, this gives the user faster load times. Video streamer Netflix has  its own CDN. This CDN is fully operational in 2015 and now they still use some other CDN's like Akamai. Reasons for having an own CDN is that the load balancing can be better improved. The CDN that Netflix has created is called Open Connect. To make their CDN possible they work together with the ISP providers. They use high performance HTTP delivery to get the content to the users. By having their own CDN they can get the video's faster to their users, because they can optimize their video stream algorithms for their own CDN. Besides this its possible that in this way the data going over the network can be decreased.

@@

There have been several cloud projects with the Raspberry Pi. \newline
One of these has been the supercomputer build by Southampton~\cite{cox:2014}. Here they build it with 64 raspberry Pi's . They used a Message Passing interface to communicate between the Raspberry Pi's. The research was done in order to see what the performance of a low-power high performance cluster is. \newline
The project of the university of Glasgow~\cite{tso:2013}. This data center has 56 Raspberry Pi's. It is build for research and education purposes for a cloud data center. They have a Hadoop running on their servers. \newline
There was another project called the Beowulf cluster~\cite{beowolf-setup}. This project was created for a PhD assignment. This cluster is build for collaboratively processing sensor data.  A Beowulf cluster is simply a collection of identical, commodity computer hardware based systems~\cite{beowolf-setup}. The big advantage of the Raspberry Pi in this project was that it was cheap and it does not need a cluster administrator to check everything that was being done.

For Video streaming there are several projects. These projects are often done by service providers such as Google with their video streamer Youtube. \newline
NGINX is busy with their Video streaming project. NGINX works together with Netflix to give users the best video watching experience. NGINX is a scalable service with high performance load balancing.  Netflix is together with NGINX busy with their project Open Connect. This project wants to improve the load balancing for video streaming over the whole network. \newline
Another service is WOWZA \cite{wowza}. WOWZA provides video streaming for a lot of companies and universities. WOWZA has done several projects for video streaming and has build with the University of Maine a content management system for their online videos. 
One mayor video stream service is of course Youtube \cite{youtube}. Youtube was once started as a project with the aim to remove the technical barriers to share videos online. Nowadays Youtube is still working hard on optimizing their video streaming service. Youtube has a online API so developers can join in and start their own projects. Youtube has its own projects for example helping mobile gamers stream their video, helping with live streaming and many more.

@@

\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{Pi2ModB1GB_-comp.jpeg}
	\caption{Raspberry Pi 2 model B}
	\label{fig:raspberry} %always place your label after your caption!
\end{figure}
\begin{table}[H]
	\centering \caption{Specifications}
	\begin{tabular}{|c|c|} \hline
		Ethernet & 100 Mbps \\ \hline
		USB & 4 x USB 2.0 \\ \hline
		Video out & HDMI 1.4 \\ \hline
		Audio & 2 x analog \\ \hline
		CPU & 900MHz quad-core ARM Cortex-A7 \\ \hline
		card slot & Micro SD  \\ \hline
	\end{tabular}
	\label{tab:Specificaties}
\end{table}
The Raspberry Pi 2 (RPi 2) only consumes 3 watts, because of this it does not need active cooling. The RPi 2 does have a limited ethernet cable of only 100 Mbps. This can be a huge drawback when large amounts of data need to be distributed over the Network. It is however very useful for a investigation in load balancing for video streaming. A small network cable offers still room for analyzing the video stream. The USB 2.0 has not a huge capabilities to share the data from a external harddisk to a user. A external Seagate harddisk demands 10 Watt when it is actively streaming videos and when it is standby it has a consumption of 0.1  Watt.

This research makes use of three Raspberry Pi's one router and one switch. All these Pi's need power, therefore there is a power supply and 3 micro USB cables. There are three sd cards needed for the Operating System, and other software, to make the RPi 2 work. A cost structure for this can be found in the appendix section~\nameref{sec:cost}.

For some benchmark test a laptop is plugged in, so that several tests can be done from the laptop.

@@

\subsection{Experiment approach}
This setup can be build by following the information in the GitHub link in section \ref{sec:software} that can be found in the appendix. More details about the setup can be found in section \ref{sec:setup}. \newline 
A RPi 2 is used for video streaming with load balancing. For the experiment Raspberry Pi's with a static ip address are needed. This so we can connect easily to them with SSH.  For the experiment we need a Raspberry Pi 2 that can make use of the Ethernet. This is because a RPi 2 is used as a webserver to make video streaming possible. This webserver will make video streaming possible. It will have a VPN to make it possible to put video's on the RPi 2. 
After this a cluster with one RPi 2 as a load balancer and the other two as a video stream Pi will be build. If such a cluster is made tests can be done and improvements can be made. For example some test in streaming in different quality with load balancing. The setup of the Raspberry Pi 2:

\begin{figure}[H]
	\centering 
	\includegraphics[width=0.5\textwidth]{raspsetup.jpg}
	\caption{Raspberry Setup}
	\label{fig:setup} %always place your label after your caption!
\end{figure}
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex] 
\item The client who wants to get the video. 
\item The router who redirects the client to the load balancer. The router will also have the video streamers in this network. If there are to much Raspberry Pi's in a network to connect to a single router a switch can be placed between them. 
\item The load balancer who redirects the client to the right video streamer.
\item The video streamer who streams the video to the client.
\item The video streamer who streams the video to the client.
\end{enumerate}
In the section \ref{sec:setup} the software is defined to create a cluster like this.

\subsection{Raspberry Pi 2 inside a data center}

>>>>

\usepackage[numbers]{natbib}
\setlength{\bibsep}{0.0pt}
\restylefloat{figure}
\restylefloat{table}
\begin{document}
\setcounter{page}{1}
\pagenumbering{arabic}
%
% --- Author Metadata here --- DO NOT REMOVE OR CHANGE 
\conferenceinfo{23$^{th}$ Twente Student Conference on IT}{June 22$^{st}$, 2015, Enschede, The Netherlands.}
\CopyrightYear{2015} % Allows default copyright year (200X) to be over-ridden - IF NEED BE.
%\crdata{0-12345-67-8/90/01}  % Allows default copyright data (0-89791-88-6/97/05) to be over-ridden - IF NEED BE.
% --- End of Author Metadata ---

@@

\keywords{Cloud computing, Raspberry Pi 2, RPi2,  micro data center, video streaming, load balancing}

\section{Introduction}
Today, most of us have data in the cloud. Despite the attention from the community, research and development of Cloud Computing services are still in its early days~\citep{tso:2013}. \newline
Cloud computing is a trend in IT that moves customers computing and data away from computer, and laptop into large data centers \citep{dikaiakos:2009}. In the future, most Internet users are expected to access Internet services over lightweight portable devices requiring a lot of network bandwidth \cite{dikaiakos:2009}. This bandwidth usage causes bottlenecks. To prevent these bottlenecks, data centers are built around the globe. These data centers can be greatly improved \cite{abrahamsson:2013,beloglazov:2010}. 
For example, these data centers require a lot of space and cooling. There are now new technologies such as the powerful ARM processor. Companies want to explore the possibilities of the Raspberry Pi 2 (RPi2), because of its ARM processor and the low price \cite{Pcextreme}. At the time of writing the RPi2 costs around 40 euro. The RPi2 is cheaper in price compared to a more expensive normal server. Some data centers offer some cloud computing using the Raspberry Pi 2 \cite{Pcextreme}. \newline
A RPi2 is a small computer and has a power usage between two and four Watt. In contrast, a normal server has a power usage between 75 and 250 Watt~\cite{Pcextreme,beloglazov2012energy}. The low power consumption of the RPi2 and its computing power could mean that it is better to use a RPi2 for specific small tasks that do not demand an entire server. For this reason, research is done on the performance of the RPi2 in a small data center. \newline
A Raspberry Pi cloud can be the small data center for the future \cite{tso:2013}. The RPi2 can be an ideal testbed for testing distributed software.

In The Netherlands one million people use Netflix \cite{volkskrant}. Netflix is a popular video streaming service that provides HD movies. To achieve this Netflix makes use of a Content Distribution Network (CDN). Netflix makes use of MPEG-DASH, a protocol for streaming over HTTP \cite{martin:2013}. The problem is that Netflix is responsible for approximately 30 \% of the peak downstream traffic in the US \cite{Adhikari:2012}. The result was that the two main providers Comcast and  AT\&T were limiting the downstream traffic of Netflix. This caused a lot of criticism, to solve this a new law for net neutrality has been made \cite{net-neutrality}. \newline
In the Netherlands, video streaming services such as Youtube and Netflix have an increasing number of users. Nowadays, the data streaming problems are increasing, this makes it interesting to do more research in data centers with video streaming. Most of the existing research in computer science is performed on expensive large servers \cite{tso:2013}. It can be very useful to see if it is possible to do research on a small computer like a RPi2.

This paper investigates how useful the RPi2 is in a data center with video streaming. To come to a conclusion the performance of the RPi2 is evaluated by doing several benchmark tests. For this the main research question is: 
\begin{center}
How well does the RPi2 perform in small scale data centers with video streaming? 
\end{center}

@@

\begin{enumerate}
	\item What are small scale data centers with video streaming and why are they used?
	\item Is the RPi2 usable for video streaming?
	\item How to fit the RPi2s into a data center considering cooling, space allocation and power?
	\item What setup does a RPi2 cloud with video streaming require?
	\item How is availability in a RPi2 cluster with video streaming affected by various load balancing techniques? 
\end{enumerate}

The five sub-questions are investigated in the next sections. Followed by a final section for conclusions and identified areas for future work.
The first two research question are a literature study into video streaming and cloud computing. The last three sub-questions go into the technical research that investigates the performance and possibility of RPi2s for cloud video streaming. This is done by building a small RPi2 cloud.

\section{Research Methods}
In the research the  aim is to build a RPi2 cloud to measure the performance. This is achieved by building a RPi2 cloud and doing benchmark test on this cloud. This research uses the Design Science method proposed by Hevner~\cite{hevner:2007}. The method is used to make an architecture for a Raspberry Pi cloud. This design science method consists of three cycles:
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{Design_science.png}
	\caption{Design Science cycles \cite{hevner:2007}}
	\label{fig:design} %always place your label after your caption!
\end{figure}
\textbf{relevance cycle} Represents the interaction with the domain. Here the problems and opportunities are identified. The identification of these problems are done in the next section. \newline 
\textbf{design cycle} The most important cycle where the design is built. This is highlighted in section \ref{sec:setupex}. \newline
\textbf{rigor cycle} Represents the interaction with the scientific community. This information can be found in section \ref{sec:back} and \ref{sec:setupex}.

\section{Background}\label{sec:back}
\subsection{Small scale computing}
Small scale cloud computing is cloud computing with smaller computation amounts than normal \cite{cox:2014}. Most customers for cloud solutions buy the services instead of buying the hardware and having to set up the servers themselves. This service is called Infrastructure as a Service (IaaS). The elasticity of computing resources in the cloud is important to provide IaaS \cite{Miettinen:2010:EEM:1863103.1863107}. Many companies have their software running on virtual machines in data centers \cite{beloglazov:2010}. The advantage is that these virtual machines are scaled dynamically to the customers need. Using the cloud makes  management easier.

Several cloud services have a huge network demand, this may cause data centers to be unreachable. Because of the privacy and reachability of a data center many companies have their own servers. These servers are not energy efficient. \newline
Netflix, a cloud service, is responsible for approximately 30\% of the peak downstream traffic in the US \cite{Adhikari:2012, computer-networking}. The data flowing through the network has different patterns. To improve this there are several data patterns possible, it can for example make a huge different in sending data in small bursts or in one large burst \cite{computer-networking}.

Power Usage Efficiency (PUE) is a important aspect in cloud computing \cite{beloglazov2012energy}. There are more lightweight portable devices coming \cite{Miettinen:2010:EEM:1863103.1863107}, these devices are energy efficient, but will demand more processes by data centers \cite{Miettinen:2010:EEM:1863103.1863107}. Nowadays, the data centers operate 12 to 20\% of the time, the other 80 \% of the time the servers inside are still using energy \cite{beloglazov2012energy, meisner2009powernap}. Therefore there is a need for Green Cloud Computing solutions that reduce the electrical power consumption and the environmental impact of data centers~\cite{beloglazov2012energy}. Microservers can be used to reduce the electricity consumption.

\subsection{Microserver}
Microservers are used in data centers \cite{microserver}. Microservers consume less electricity and have enough computer power for small tasks, therefore they can be used for high traffic usage tasks. Besides this they are small and are easier replaced compared to a normal server, by using them data centers can be more dynamic \cite{microserver}. The ability to add and remove microservers more easily can save a lot of energy. \newline
A microserver often has a different processor compared to a normal server. The processors are often microprocessors like ARM processors \cite{microserver}. There are three big processor companies in the world: ARM, Intel and AMD. The Intel and AMD processors are better at complex tasks, while the ARM processor is better at small tasks \cite{microserver}. In the data center of the future more dedicated microservers will be used, they are better at specialized tasks and are more energy efficient \cite{microserver}.

\subsection{Video streaming}\label{sec:stream}
HTTP is used for video streaming, HTTP sends frames to the client and in the frames is the video stored \cite{computer-networking}. If the send rate is large enough the video will play without any delays, but if the send rate is too low then the video playback will alternate between periods of continuous playback and periods of freezing~\cite{computer-networking}. \newline 
To make video streaming possible the client will connect with video streaming service by its load balancer. The load balancer will then locate the videos. To show how the processes and operators operate with each other, the UML sequence diagram in figure \ref{fig:stream} is used \cite{Bernardi:2002:USD:584369.584376}. 
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{VideoStreaming.jpg}
	\caption{Video stream}
	\label{fig:stream} %always place your label after your caption!
\end{figure}

In the diagram there is a loop for the video streaming, here the client will keep requesting more parts of the videos. Depending on the algorithm being used the amount of data requested by the client can fluctuate. For streaming a video HTTP streaming is often used. If the client receives too much data its buffer gets full. To prevent throwing data away the server then stops sending.

There is a important streaming technique called adaptive streaming \cite{computer-networking}. This streaming algorithm is often used for video on demand (VOD) \cite{computer-networking}. It improves the experience for the end-user by adapting the video quality dynamically to the viewers network condition \cite{ffmpeg}. This techniques takes into account the clients CPU possibility and its network possibilities. It delivers small sized chunks at multiple bitrates. This techniques is used for HTTP Live Streaming (HLS) \cite{jwplayer}. HLS makes use of RTMP streaming and HTTP. HLS codifies video in the H.264 format.\newline
There is another special kind of HTTP streaming called DASH (Dynamic Adaptive Streaming over HTTP)~\cite{computer-networking}. In DASH the video is encoded into several different versions. If the bandwidth is high, then the client selects chunks from a high-rate version and the other way around~\cite{computer-networking}. The possibility to switch between versions of different bitrates is in video streaming important for user experience. To make a smooth transition between these versions possible there are intermediate versions so that the client will see the change in video quality less suddenly.

\subsection{Video streaming and load balancing}
For video streaming several load balancing techniques are used. To choose a technique the following aspects are taken into account performance, reliability and features. There can be a non-uniform demand for different movies. One solution to overcome this problem would be to replicate the most popular movies, but this is expensive in terms of storage space required for this. A dynamic replication can be used to solve the load demand~\cite{dan1996load}. Parts of a movie are moved to less used storage devices, so that these can stream the movie. The allocation of these movies only happens when there is a too high increase on one specific storage device. In this way load imbalances can be prevented \cite{dan1996load}. \newline
Load balancers are generally grouped in two layers. Layer 4, the transport layer and layer 7 the application layer \cite{computer-networking}. Layer 7 load balancers distribute requests based upon data in the application layer. Here below are two protocols and a language that is used:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex]
	\item HTTP, is the hypertext transfer protocol. In section \ref{sec:stream} usages of this protocol are mentioned. A lot of video streaming services use HTTP streaming \cite{Adhikari:2012}.  
	\item RTMP is a real time messaging protocol ~\cite{rtmp}. It is developed for streaming data between a player and a server. It is encapsulated in HTTP to traverse firewalls. It gives more video player options compared to normal HTTP, thus giving the user a better experience. A disadvantage compared to normal HTTP is that it is sensitive for data spikes. These data spikes can result in a overload of the buffer which results in a empty buffer, this can lead to a stop in the video playing. 
	\item Synchronized Multimedia Integration Language (SMIL) describes multimedia presentations \cite{smil}. SMIL refers to media objects by URLs, allowing them to be shared between presentations and stored on different servers for load balancing. 
\end{enumerate}

@@

Users want on demand video streaming, because they do not want to store the data themselves, and want a wide choice of different videos on demand. Netflix is a on demand video streaming service that makes HD movies watching possible. For this Netflix uses a content distribution network (CDN). Netflix uses amazon its AWS, simpleDB, S3 and Cassandra for file storage \cite{Adhikari:2012}. Netflix makes use of MPEG-DASH, a protocol that makes streaming over HTTP possible. To make this possible Netflix does several things in the cloud.

@@

\subsubsection{Video stream CDN}
A CDN uses small data centers, because users want their data as nearby as possible  In this way the latency is decreased and there is a shorter connection distance, this gives the user faster load times. Video streamer Netflix has  its own CDN \cite{netflix}. Reasons for having an own CDN is that the load balancing can be improved. The CDN that Netflix has created is called Open Connect. Netflix uses high performance HTTP delivery to get the content to the users. Data going over the network can be decreased and videos get faster to users with an own CDN, because video stream algorithms are better analyzed.

@@

There have been several cloud projects with the Raspberry Pi. \newline
One was the supercomputer built by Southampton~\cite{cox:2014}. Here a cluster was built with 64 raspberry Pi's. A Message Passing interface was used to communicate between the Raspberry Pi's. The research was done in order to see what the performance of a low-power high performance cluster was. \newline
The project of the university of Glasgow~\cite{tso:2013}. This data center has 56 Raspberry Pi's. It was built for research and education purposes for a cloud data center. Hadoop was tested on their servers. \newline
Another project called the Beowulf cluster~\cite{beowolf-setup}. This project was created for a PhD assignment. This cluster was built for collaboratively processing sensor data.  A Beowulf cluster is simply a collection of identical, commodity computer hardware based systems~\cite{beowolf-setup}. The big advantage of the Raspberry Pi in this project was that it was cheap and it does not need a cluster administrator to check everything that was being done.

For Video streaming there are several projects. These projects are often done by service providers. \newline
Netflix works together with NGINX on Netflix his project Open Connect. With this cooperation they want to give  users the best video watching experience. This project wants to improve the load balancing for video streaming over the whole network. \newline
Another service is WOWZA \cite{wowza}. WOWZA provides video streaming for a lot of companies and universities. WOWZA has done several projects for video streaming and has built with the University of Maine a content management system for their online videos. \newline 
One mayor video stream service is of course Youtube \cite{youtube}. Youtube was once started as a project with the aim to remove the technical barriers to share videos online. Nowadays Youtube is still working hard on optimizing their video streaming service. Youtube has a online API, so developers can join in and start their own projects. Youtube has its own projects for example helping mobile gamers stream their video and helping with live streaming.

@@

\begin{figure}[H]
	\centering 
	\includegraphics[width=0.4\textwidth]{Pi2ModB1GB_-comp.jpeg}
	\caption{Raspberry Pi 2 model B \cite{raspberry-pi}}
	\label{fig:raspberry} %always place your label after your caption!
\end{figure}
\begin{table}[H]
	\centering \caption{Specifications \cite{raspberry-pi}}
	\begin{tabular}{|c|c|} \hline
		Ethernet & 100 Mbps \\ \hline
		USB & 4 x USB 2.0 \\ \hline
		Video out & HDMI 1.4 \\ \hline
		Audio & 2 x analog \\ \hline
		CPU & 900MHz quad-core ARM Cortex-A7 \\ \hline
		card slot & Micro SD  \\ \hline
	\end{tabular}
	\label{tab:Specificaties}
\end{table}
The RPi2 only consumes 3 Watt, because of this it does not need active cooling. The RPi2 has a limited ethernet interface of only 100 Mbps. The network cable offers room for analyzing the video stream. The USB 2.0 is used to share the data from a external harddisk to a user. A external Seagate harddisk demands 10 Watt when it is actively streaming videos and when it is standby it has a consumption of 0.1  Watt.

This research makes use of three Raspberry Pi's one router and one switch. All these Pi's need power, therefore there is a power supply and 3 micro USB cables. There are three sd cards needed for the Operating System, and other software, to make the RPi2 work. A cost structure for this can be found in the appendix section~\nameref{sec:cost}.

For some benchmark tests a laptop is plugged in, so that several tests can be done from the laptop.

@@

\subsection{Experiment approach}
This setup is built by following the information in the GitHub link in section \ref{sec:software}, that can be found in the appendix. More details about the setup can be found in section \ref{sec:setup}. \newline 
A RPi2 is used for video streaming with load balancing. For the experiment we need a Raspberry Pi 2 that can make use of the Internet. This is because a RPi2 is used as a webserver to make video streaming possible. It has a VPN to make it possible to put video's on the RPi2. 
A cluster with one RPi2 as a load balancer and the other two as a video stream Pi is built. If such a cluster is made tests can be done and improvements can be made. For example streaming tests with different quality videos and load balancing. The test are found in section \ref{sec:test}. Below the setup design and legend for the RPi2 are given:
\begin{enumerate}[topsep=0pt,itemsep=-1ex,partopsep=1ex,parsep=1ex] 
	\item The client who wants to get the video. 
	\item The router who redirects the client to the load balancer. The router will also have the video streamers in this network. If there are to much Raspberry Pi's in a network to connect to a single router a switch can be placed between them. 
	\item The load balancer who redirects the client to the right video streamer.
	\item The video streamer who streams the video to the client.
	\item The video streamer who streams the video to the client.
\end{enumerate}
For the research number three to five are Raspberry Pi's.
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.5\textwidth]{raspsetup.jpg}
	\caption{Raspberry Setup}
	\label{fig:setup} %always place your label after your caption!
\end{figure}

\subsection{RPi2 inside a data center}

In this part the research focuses on the possibilities of the RPi2 in a data center. \newline
It is possible to fit a lot of RPi2s inside a data center. PCextreme has placed 500 Pi's inside a single data rack \cite{Pcextreme}. For this special designed boards are needed. With a special board RPi2 can be rebooted and turned off from a distance \cite{Pcextreme}. By having a microserver like a RPi2 in a data center the hardware can be owned by only the customer, because its better payable than a normal server. In a data center there are the advantages of the high speed Internet. By owning the hardware the customer will have more control over the data in the privacy and reliability aspect.\newline In this research there are 3 Pi's and these fit easily inside the rack. The cooling of the data rack is sufficient for all these RPi2s. The power consumption of the RPi2 is low as seen in table \ref{tab:cpu}. The RPi2 is easy to replace and it can easily be turned on and off, this makes it suitable inside a server rack. A problem that will exist with the RPi2 inside the data center is that every time a Pi's have troubles a help team inside the data center has to fix this for the customer.

In this project a cluster of RPi2s is built to test its possibilities. The setup for this project is displayed below:
\begin{figure}[H]
	\centering 
	\includegraphics[width=0.45\textwidth]{setup.jpg}
	\caption{Project Setup}
	\label{fig:projectsetup} %always place your label after your caption!
\end{figure}
The numbers in front in figure \ref{fig:projectsetup} point to the RPi2s, these numbers correspond with the numbers in figure \ref{fig:setup}. The number in the back shows the little router that is behind the power supply. In this picture there is a extra switch to connect all the Raspberry Pi's. \newline 
This setup can be placed inside of a data center. The switch and the router will then be removed and the rest will be fit into the server rack.
